---
title: "ChiSquared"
output: pdf_document
date: "2024-03-02"
---
## Experience with calculating chi-squared and p-values.

```{r}
classes_observe <- matrix(c(41, 10, 32, 2, 4, 12, 11, 31, 5, 10, 15, 30), ncol = 2, byrow = TRUE,
                   dimnames = list(c("Math (F-B)", "Math (A)", "English (F-B)", "English (A)", "Science (F-B)", "Science (A)"),
                                   c("Studied", "No Studying")))
classes_observe
```
```{r}
observed <- matrix(c(50,32,58,63), ncol = 2, byrow = TRUE,
                   dimnames = list(c("F-B",  "A"),
                                   c("Studied", "No Studying")))
observed
```

```{r}
# (a) Calculate marginal frequencies
row_marginals <- rowSums(observed)
col_marginals <- colSums(observed)

# Print row and column marginals
print("Row Marginals:")
print(row_marginals)

print("Column Marginals:")
print(col_marginals)
```
```{r}
# (b) Calculate row proportions
row_proportions <- observed / sum(row_marginals)
row_proportions
```
```{r}
# (c) Calculate column proportions
col_proportions <- observed / sum(col_marginals)
col_proportions
```
```{r}
# (d) Calculate expected values
expected <- outer(row_marginals, col_marginals) / sum(observed)
expected
```
```{r}
# (e) Calculate residuals
residuals <- (observed - expected) / sqrt(expected)
residuals
```
```{r}
# (f) Calculate chi-squared values for each residual
chi_squared_values <- residuals^2
chi_squared_values
```
```{r}
# (g) Calculate total chi-square measure
total_chi_squared <- sum(chi_squared_values)
total_chi_squared
```


```{r}
# (h) Find the p-value
degrees_of_freedom <- (nrow(observed) - 1) * (ncol(observed) - 1)
p_value <- 1 - pchisq(total_chi_squared, df = degrees_of_freedom)
p_value
```


```{r}
# (i) Conclusion
confidence_level <- 0.85
if (p_value < 1 - confidence_level) {
  cat("There is a significant association between studying and getting an A across all subjects at 85% confidence level.\n")
} else {
  cat("There is no significant association between studying and getting an A across all subjects at 85% confidence level.\n")
}

# Check if the results would be significant by a more typical publishable standard (95% confidence)
publishable_confidence_level <- 0.95
if (p_value < 1 - publishable_confidence_level) {
  cat("The results are significant by a more typical publishable standard (95% confidence).\n")
} else {
  cat("The results are not significant by a more typical publishable standard (95% confidence).\n")
}
```
At the 85% confidence level, I compare the p-value to the significance level (α). For an 85% confidence level, α would be 1−0.85=0.15.
Conclusion for 85% Confidence Level: The p-value (0.067655213) is less than the 15% significance level (0.15). Since the p-value is lower than the significance level, we reject the null hypothesis. There is evidence of a significant association between studying and getting an A across all subjects at the 85% confidence level.
At the 95% confidence level, I compare the p-value to the significance level (α). For an 95% confidence level, α would be 1−0.95=0.05.
Conclusion at 95% Confidence Level: The p-value (0.067655213) is greater than the 5% significance level (0.05). Since the p-value is higher than the significance level, we fail to reject the null hypothesis. There is not enough evidence to claim a significant association between studying and getting an A across all subjects at the 95% confidence level.
Result: At 85% Confidence Level: Significant association, but at 95% Confidence Level: Not a significant association.

